---
title: "Military Combat Deaths Compared with Military Suicides"
output: html_document
date: 'April 3, 2022'
---

``{r Packages}
install.packages(c('gapminder','ggplot2','gganimate','gifski', 'tidyr', 'ggrepel'))
```

```{r Library}
library(gganimate)
library(ggplot2)
library(gifski)
library(tidyr)
library(ggrepel)
theme_set(theme_bw())
```


1) First, let's make our data set. 
```{r making data}
MilStats <- data.frame(suicides = c(153, 155, 174, 190, 197, 182, 213, 211, 268, 309, 295, 301, 525, 475, 444, 478, 478),
                  combat_deaths = c(0, 0, 0, 312, 735, 739, 769, 847, 352, 346, 456, 394, 259, 0, 0, 0, 0),
                  year = c(2000, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016))
```

2) Save the dataset as csv
```{r Df to CSV}
library(readr)
write_csv(MilStats, "~/Documents/GitHub/Portfolio3-Mil_Combat_Data/MilStatsExcelTest.csv")
```

3) Take a look at our static, non-animated plot
```{r Scatterplot}
ggplot(data = MilStats, aes(x= year, y = combat_deaths, suicides)) +
   geom_point(alpha = 0.7, show.legend = FALSE) +
   labs(title = 'Year: 2000-2016', x = 'Year', y = 'Number of Deaths / Suicides')
```


Let's enter one variable at a time, starting with suicides.
```{r MultipleVars1}
plot(MilStats$year,                    # Draw first time series
     MilStats$suicides,
     type = "l",
     col = 2,
     ylim = c(0, 1000),
     xlab = "Year",
     ylab = "Values")
legend("topright",                     # Add legend to plot
       c("suicides"),
       lty = 1,
       col = 2:4)
```

Next, let's plot our second variable, combat deaths, with a different color.
```{r MultipleVars2}
plot(MilStats$year,                    # Draw second time series
     MilStats$combat_deaths,
     type = "l",
     col = 3,
     ylim = c(0, 1000),
     xlab = "Year",
     ylab = "Values")
legend("topright",                     # Add legend to plot
       c("combat_deaths"),
       lty = 1,
       col = 3:4)
```


Next, we combine the two. 
```{r Combine}
plot(MilStats$year,                    
     MilStats$suicides,
     type = "l",
     col = 2,
     ylim = c(0, 1000),
     xlab = "Year",
     ylab = "Values")
lines(MilStats$year,                    
     MilStats$combat_deaths,
     type = "l",
     col = 3)
legend("topright",                     
       c("combat_deaths", "suicides"),
       lty = 1,
       col = 2:3)
```


Finally, let's animate this!

``{r Animate}
plot(data=MilStats$year,                    
     MilStats$suicides,
     type = "l",
     col = 2,
     ylim = c(0, 1000),
     xlab = "Year",
     ylab = "Values")
lines(MilStats$year,                    
     MilStats$combat_deaths,
     type = "l",
     col = 3)
legend("topright",                     
       c("combat_deaths", "suicides"),
       lty = 1,
       col = 2:3)
labs(title = 'Year: {frame_time}', x = 'Year', y = 'Deaths due to Combat and Suicide') +
             transition_time(year) +
             ease_aes('linear')
animate(MilStats)
```

This gave me the error "animation of data.frame objects not supported"

So, I hunted for a different code. Dutch data scientist Joαchim Goedhαrt has a terrific script that provides explanations for the code:
https://github.com/JoachimGoedhart/Animate-Labeled-TimeSeries

I edited my Excel file by hand, so that the relationship between my Excel formatting and my code would match Joαchim Goedhαrt. This made for an incredibly easy transition. Lesson learned: finding the code can be harder than making it run, at time. Never quit!

I read my new Excel doc into a dataframe
```{r Read data from CSV into a dataframe}
library(readr)
dfMilStatsTest <-  read.csv("~/Documents/GitHub/Portfolio3-Mil_Combat_Data/MilStatsExcel.csv")
```

Next, for the coup-de-main: the animation code (all comments following # are fully credited to Mr. Goedhαrt)

```{r Define aes}
ggplot(dfMilStatsTest, aes(Time, Value, group = Object, color = Object)) + 
  geom_line(size=1) +   
  geom_point(size = 2) + 
  geom_label_repel(aes(x = Time, y=Value, label = Object, fill = Object), hjust = 0, direction    = "y", nudge_x = 20,
                   fontface = 'bold',
                   color = 'white', size=6,
                   segment.color = 'grey50',
                   segment.size = 0.5) +
  transition_reveal(Time) 
```

Even after successfully creating the animation, when I knitted, the code above gave me an error in Line 133, where the # comment is. So I have deleted those comments in the code and pasted them below.
``{r Code with # comments}
#Define the aesthetics for the plot; x=Time, y=Value and each object is grouped and given a unique color
ggplot(MilStatsExcel, aes(Time, Value, group = Object, color = Object)) + 
  geom_line(size=1) +   
  #Show a dot for the new datapoint that is added in a new frame (works well in the animation)
  geom_point(size = 2) + 
  #Add labels, geom_label_repel from the ggrepel package is used to avoid overlap between the labels
  geom_label_repel(aes(x = Time, y=Value, label = Object, fill = Object), hjust = 0, direction    = "y", nudge_x = 20,
                   fontface = 'bold',
                   color = 'white', size=6,
                   segment.color = 'grey50',
                   segment.size = 0.5) +
   #This is where the magic happens       
  transition_reveal(Time) 
``


Last, remember to save plot to your file!
```{r Save plot to file}
ggsave('MilStats.png', width=8, height=8)
```
That did not work, so I just saved manually!


